---
title: "lab_01"
author: "Bex Heimbrock"
date: "2025-09-10"
output: html_document
---

PUT YOUR NAME AND DATE ON THIS ASSIGNMENT



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your class reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the UMD courses data.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
umd_courses <- read_rds("lab_01.rds")
```

## Answer questions

**Q1** Which department has the highest mean (average) number of seats for its classes? You need to show the department with the highest average first. Which one is the most surprising/interesting to you? **A1** The department with the highest average seats is First-Year Innovation and Research Experience. This is not surprising if it's a gen-ed course for all first-years. I find it interesting, and sad, that the African American studies dept only has, on avergae, 44 seats. Come on, humanities!

```{r}
umd_courses |>
  group_by(term, department) |>
  summarise(
    count_classes = n(),
    total_seats = sum(seats),
    mean_seats = mean(seats),
    median_seats = median(seats)
  ) |>
  arrange(desc(mean_seats))
```

**Q2** Find out the greatest number of times any individual class has been offered in our dataset. Your result should sort to show the greatest number of classes first. If several classes are tied for the most offered, describe how many classes are ranked in the top category for offerings.
**A2** Roughly 145 classes are ranked in the top category for offerings. 

```{r}
umd_courses |>
  group_by(id) |>
  summarise(
    count_classes = n()
  ) |>
  arrange(desc(count_classes))
```

**Q3** Count the number of courses for each instructor and department. Which instructor has taught the most? HINT: To filter out blank and missing values from a column, drop_na() is your friend.
**A3** Keith Yearwood

```{r}
umd_courses %>% 
    drop_na(instructors) %>% 
    group_by(instructors, department) %>% 
    summarise(
        count_classes = n()
    ) %>% 
    arrange(desc(count_classes))
```

**Q4** Given the results for Q3, what would have made it easier to find the answer? You don't have to know *how* to accomplish this; I want you to describe in words what would make the answer easier to see. How confident are you that the answer you got is the right one, and why? **A4** Knowing how to code would have made it easier. But also, I think the answer would be easier to see the answer if there were less columns - so filtering out more things? I'm very confident that the answer I got is the right one because Professor Wells helped me with my code.

**Q5** Take one of the questions you listed in the final question of the pre_lab and try to answer it here in code. If you don't know how to fully answer it, write what code you can and describe in words what else you would need to do. If you got an answer, is it interesting? **A5** How do the seat number trends change by year (per department)? We were able to write code that filters by year to find a trend for that year. What I want is to find yearly trends per department. Still not 100 percent sure how to find that...

```{r}
umd_courses %>% 
  group_by(department, seats, term) %>% 
  filter(term==201801) %>% 
  summarise( 
    count_classes = n()
    ) 
```
